A Comparison of Parallel Graph Processing Platforms

### 2017-01-26 papers/gabb17/gabb17-pollard.tex
Updated gabb17 paper.

\begin{document}
\maketitle
\begin{abstract}
This is a survey of existing graph analytics frameworks and an analysis of their performance.
\end{abstract}

@@

[Talk about WHY we care about graph processing, and parallel in particular. Talk about what parallel graph processing is.]

Our research is motivated by the current state of parallel graph processing. The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems \cite{Doekemeijer:2015:GPFSurvey}. These systems operate with a wide range of parallelism paradigms such as GPU [Medusa], shared memory [pretty much everything], a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, distributed memory with MPI [cite], domain-specific languages [Green Marl], as well as novel communication schemes [activemessages, sockets].

Since 2014, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo} [name some more]. At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' incredibly difficult. There has even been a propagation of so-called ``reference implementations'' which provide implementations of the most common graph algorithms [GAP, GraphBIG, Galois]. Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial. To quote Andrew Tanenbaum, ``The nice thing about standards is that you have so many to choose from.''

[This paper aims to simplify the selection process for graph processing.]

Another issue with the parallel graph processing library designers is the lack of comprehensive comparison. One possible reason for this is the considerable effort of getting each library and package to work, satisfying dependencies, and ensuring the data is in the correct format for each system. Beyond this, each system may have a different method of measuring performance. Thus, one of the contributions of this paper is to provide a ``level playing field'' for each graph processing system. Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but has not seen widespread adoption---only eight platforms are supported\footnote{According to \url{https://github.com/tudelft-atlarge}}. [TODO: Say more about why this is justified instead of just using Graphalytics]

\section{Algorithms}
The canonical performance leaderboard for paralell graph processing is the Graph500 \cite{Murphy:2010:Graph500}. The advantage of the Graph500 is it provides standardized measurement requirements and dataset generation. The primary drawback with using reference implementations for the Graph500 is the standard only supports a single algorithm: breadth first search (BFS).

This report attempts to add similar rigor to other graph algorithms. This is done by borrowing heavily from the Graph500 specification; we use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19,
C = 0.19,$ and $D = 1-(A+B+C) = 0.05$. The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}. Some clarifications are in order: for the first kernel, the time to read in the graph from disk is not considered. Furthermore, these data structures need only be created once and BFS is performed on 64 roots. The second kernel creates a tree with distances from the root, but need only label the existing data structure (and not output the entire graph, for example).

One straightforward extension is computing the Single-Source Shortest Paths algorithm (SSSP). We use the same graph and the same source vertices as in BFS.

\section{The Existing Systems}
This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}.

Parallel Boost Graph Library \cite{Gregor:2005:PBGL} and PowerGraph \cite{Gonzalez:2012:Powergraph} implementations are also provided because of their popularity and the availability of high quality reference implementations.

@@

\begin{table}[htb]
	\centering
	\begin{tabular}{l|c|c|c}
		& BFS & SSSP & PR \\ \hline
		Graph500 & & & \\ \hline
		PBGL  & & & \\ \hline
		GAP  & & & \\ \hline
		GraphBIG  & & & \\ \hline
		Galois  & & & \\ \hline
		PowerGraph  & & &
	\end{tabular}
	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
	\label{tab:reportcard}
\end{table}

Graphalytics without the use of the Granula plugin produces performance measurement in two forms: runtime in seconds and traversed edges per second.

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}
In this paper we analyze the performance of four parallel graph processing systems to determine their performance and scalability using synthetic and real-world datasets. We perform analysis for three algorithms: breadth first serach, single source shortest paths, and PageRank. This paper examines previously overlooked aspects of parallel programming performance such as file I/O in addition to a more detailed performance analysis.
\end{abstract}

@@

Our research is motivated by the current state of parallel graph processing. The most comprehensive survey, released in 2014, identified and taxonomized over 80 different parallel graph processing systems \cite{Doekemeijer:2015:GPFSurvey}. These systems operate with a wide range of parallelism paradigms and target architectures such as GPU \cite{Zhong:2014:Medusa}, \cite{Kang:2009:Pegasus}, shared memory [pretty much everything], a combination of CPU and GPU \cite{Gharaibeh:2012:Totem}, distributed database querying, \cite{Rodriguez:2015:Gremlin}, distributed filesystem based approaches \cite{Xin:2013:GraphX}, distributed memory with MPI \cite{Hong:2015:PGX}, domain-specific languages \cite{Hong:2012:GreenMarl}, as well as novel communication schemes \cite{Edmonds:2013:ActiveMessages}.

Since 2014, the problem has compounded with the addition of even more proprietary and open source projects such as \cite{Cheramangalath:2015:Falcon}, \cite{Perez:2015:Ringo} [name some more]. At the outset, this plethora of choices makes the question, ``which system is the best for my problem?'' daunting. There has even been a propagation of so-called ``reference implementations'' which  implement the most common graph algorithms [cite GAP, GraphBIG, Galois]. Thus, even selecting a standard and a benchmark over which to compare various implementations is nontrivial. To quote Andrew Tanenbaum, ``The nice thing about standards is that you have so many to choose from.''

Another issue among parallel graph processing systems is the lack of comprehensive comparisons. One possible reason for this is the considerable effort involved in getting each system to run: satisfying dependencies and ensuring correctly formatted data the data are nontrivial for most tasks. Beyond this, each system may have a different method of measuring performance. Thus, one of the contributions of this paper is to provide a ``level playing field'' for each graph processing system. Graphalytics \cite{Capota:2015:Graphalytics} attempts to remedy this but
[state the limitations of graphalytics]
%has not seen widespread adoption---only eight platforms are supported\footnote{According to \url{https://github.com/tudelft-atlarge}}. [TODO: Say more about why this is justified instead of just using Graphalytics]

\section{Algorithms}
The canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}. The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation. The primary drawback with the Graph500 is it measures a single algorithm: breadth first search (BFS).

This report attempts to add similar rigor to other graph algorithms. This is done by borrowing heavily from the Graph500 specification; we use a Kronecker graph \cite{Leskovec:2010:Kronecker} with initial parameters of $A = 0.57, B = 0.19,
C = 0.19,$ and $D = 1-(A+B+C) = 0.05$. The Graph500 Benchmark 1 (``Search'') is concerned with two kernels: the creation of a graph data structure from an edge list stored in RAM and the BFS\footnote{For a complete specification, see \url{http://graph500.org/specifications}}. Some clarifications are in order: for the first kernel, the time to read the graph from disk is not considered. Furthermore, these data structures need only be created once and BFS is performed on 64 roots. The second kernel creates a tree with distances from the root, but need only label the existing data structure (and not output the entire graph, for example).

One straightforward extension is computing the Single-Source Shortest Paths algorithm (SSSP). We use the same graph and the same source vertices as in BFS. We use SSSP and PageRank in this paper because of their simplicity and ubiquity.

\section{The Existing Systems}
This report explores a small sample of the existing graph processing platforms with a focus on the so-called ``reference implementations:'' the Graph500\footnote{We used version 2.1.4 from \url{http://graph500.org/referencecode}.}, GAP \cite{Beamer:2015:GAPBench}, and GraphBIG \cite{Nai:2015:Graphbig}. GraphMat is also used because it has been cited as the most performant [cite]. Our target is shared memory CPU processing.

Other popular libraries such as the Parallel Boost Graph Library \cite{Gregor:2005:PBGL} %and PowerGraph \cite{Gonzalez:2012:Powergraph} 
are not considered here because they are a PITA to get working and the authors never responded to my emails.

@@

%\begin{table}[htb]
%	\centering
%	\begin{tabular}{l|c|c|c}
%		& BFS & SSSP & PR \\ \hline
%		Graph500 & & & \\ \hline
%		PBGL  & & & \\ \hline
%		GAP  & & & \\ \hline
%		GraphBIG  & & & \\ \hline
%		Galois  & & & \\ \hline
%		PowerGraph  & & &
%	\end{tabular}
%	\caption{Performance. Note that implementations of every algorithm are not availabe for every platform.}
%	\label{tab:reportcard}
%\end{table}

@@

\begin{table}[!htb]
	\centering
	\begin{tabular}{l|c|c|c}
		System & Load Graph & Make Data Structure & Run BFS \\ \hline
		Graph500 & 0.3474 & 0.3971 & 0.003380 \\
		GraphBIG & 37.22  & N/A    & 0.1528 \\
		GraphMat & 0.1511 & 1.101  & 0.1031 \\
		GAP      & 2.351  & 0.1935 & 0.001393 \\
	\end{tabular}
	\caption{Times are in seconds. The Graph500 generates the graph instead of loading it into a file. GraphBIG builds the graph and reads in the file simultaneously. These results were averaged across 64 roots. [TODO: Why does the first root often have degree 0?]}
	\label{tab:fileload}
\end{table}

[Compare lines of code for implementation]
